The Galactic Midplane Is Not a Plane: Implications for Dynamical Analysis with {\em Gaia} Data and Beyond

### 2019-05-04 paper/ms.tex
Update Jz hist plot
Update panel referencing

We consider that the immortal observer makes a measurement every $\Myr$. At each measurement, we compute the actions as outlined above using the true potential but by specifiying the star's starting position using the systematically offset coordinate system.
Essentially we are shifting the
integrated orbit and then reintegrating at each point along the shifted orbit,
where the shifted orbit is the observed orbit. The erroneous computed actions
for each phase-space starting point are shown for the first $\Gyr$ of the
orbit in the {\em upper} panels
Figure~\ref{fig:one_orbit_wrong_ref}.\footnote{Occasionally the numerical
scheme fails and very large actions are reported by \texttt{gala}---we perform
a $5\sigma$ clip on each action to exclude such orbits, but this only excludes
a total of $5$ orbits out of the $1000$ considered for
Figure~\ref{fig:one_orbit_wrong_ref}. Some numerical artifacts remain, but the
vast majority of orbits are computed properly.}

@@

For a systematic offset in $z$ ({\em upper} panels), the
95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $2.2$ and
$6.2\,\actunit$ for $J_R$ and $J_z$, respectively. These are fractional errors
of $5.7\,\%$ and $86\,\%$, respectively. The error induced in $J_{\phi}$ is
negligible, as expected since $J_{\phi}$ only depends on the $x$- and
$y$-components of the position and velocity of the stars.\footnote{In practice,
however, $J_{\phi}$ is computed as part of the torus-fitting method.} It is worth
pointing out that a $100\,\pc$ error in an orbit with
$z_{\text{max}}=850\,\pc$ --- a $12\%$ error
--- induced an $86\%$ error in the computation of $J_z$.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{fig/schmactions_one_orbit.pdf}
\end{center}
\caption{The artificial phase-dependence in the observed actions induced by an
error in the Galactocentric coordinate system. We consider here the thick-disk
orbit, which has actions of $(J_R, J_{\phi}, J_z) = (37.9, -1520,
7.0)\,\actunit$ and $z_{\text{max}}=850\,\pc$ (see Table~\ref{tab:orbits}). We
integrate the orbit according to the procedure laid out in
Section~\ref{ssec:action_comp}. Then, we subtract $100\,\pc$ from the $z$
value ({\em upper panels}) or the $x$ value ({\em lower panels}) of each
position in the orbit, corresponding to an erroneous observer assuming a
midplane ({\em upper}) or solar radius ({\em lower}) that is off by
$100\,\pc$. We then allow our (immortal) observer to observe the orbit over
$1\,\Gyr$ and perform the same orbit integration procedure at each timestep,
and report the values of the actions. The computation of $J_{\phi}$ is
pristine to errors in $z$, with only numerical artifacts remaining. Only small
errors are induced in $J_R$, with the middle $90\%$ of values over the $\Gyr$
being within $\sim8\%$ of the true $J_R$. As expected, large errors are
induced in $J_z$ with a $100\,\pc$ offset in $z$, with the middle $90\%$ of
values being within $\sim43\%$ of the true $J_z$. The $x$ offset induces
uncertainties in $J_R$ and $J_{\phi}$ of $\sim21\%$ and $\sim3\%$,
respectively. A $\sim3\%$ error in $J_z$ is also induced.}
\label{fig:one_orbit_wrong_ref}
\end{figure*}

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{fig/schmactions_Jz_hist.pdf}
\end{center}
\caption{A histogram of the observed values in $J_z$ for the thick-disk orbit
assuming a $z$ offset of $100\,\pc$. One can see that if the observed $z$
values have a bias (from e.g. an incorrectly computed midplane), then the
induced error distribution in $J_z$ is decidedly non-Gaussian. Therefore, any
sort of error propagation must take this into account. Intuition for the shape
of each panel is given in the text. We also plot one half the 95\uth{}
percentile minus the 5\uth{} percentile of each distribution as a horizontal
arrow anchored on the true $J_z$ value. We call this $\Delta J_z$ and will use
it (along with the similarly defined $\Delta J_R$ and $\Delta J_{\phi}$) to
empirically describe the error distribution. We see that $\Delta J_z$ roughly
corresponds to the distance from the true $J_z$ value to one of the modes of
the error distribution.}
\label{fig:Jz_hist}
\end{figure}

In Figure~\ref{fig:Jz_hist} we plot a histogram of the observed values of
$J_z$ for the thick-disk orbit ({\em upper panel}) and thin-disk orbit ({\em
lower panel}) assuming a $z$ offset of $100\,\pc$ (i.e. as in the {\em upper
right} panel of Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$
is plotted as a vertical dashed line. Here we see that the errors in $J_z$
induced by a systematic offset in $z$ are non-Gaussian and bimodal.
Furthermore, neither of the modes lie at zero. In other words, neither of the
modes of the distribution of observed $J_z$ values are centered on the true
$J_z$ (i.e. the $J_z$ computed with no offset in $z$). In the case of the
thin-disk orbit ({\em lower panel}), we see that, in addition to the prior
complications, the error distribution is not even centered on null. This comes
about when the midplane error is $\gtrsim z_{\text{max}}$, where
$z_{\text{max}}$ is the maximum height of the orbit (equivalent to $A_z$ in
Section~\ref{ssec:epi_action}).

We now describe the intuition behind the shape of Figure~\ref{fig:Jz_hist}.
Consider first the thick-disk orbit ({\em top} panel), where the offset in $z$
is much less than $z_{\text{max}}$ of the orbit. The peaks in the distribution
correspond to the turning points of the orbit (or points of maximum vertical
amplitude), where $v_z \sim 0$ and where the star is on most of it orbit.
This is why the distribution is peaked on these values. For the thin-disk
orbit ({\em bottom} panel), the offset in $z$ is comparable to
$z_{\text{max}}$. Now, there will be some points in the orbit where $v_z = 0$
and $z=0$ (in the observed, erroneous coordinate system). At these points, the
computed $J_z$ will vanish. The asymmetry and systematic offset then comes
about because of the constraint that $J_z \geq 0$.\footnote{This is similar to
arguments in cosmology for why gravity produces non-Gaussianity in the density
field, since the density cannot become negative but it can grow arbitrarily
large.}

@@

The top row of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread induced
in each action for an offset in the $z$-component (i.e. the Galactic
midplane). In the {\em bottom row} we consider offsets in the $x$ component
(i.e. the Solar radius). The {\em left}, {\em middle}, and {\em right} columns
show the fractional spread in the values computed for $J_z$, $J_{\phi}$, and
$J_R$, respectively. We compute the fractional spread $\Delta J_i/J_i$, where
$\Delta J_i$ is defined in the previous paragraph and $J_i$ is the action
value computed in the true coordinate system, over the course of the first
$\Gyr$.

In the {\em upper middle} panel of Figure~\ref{fig:many_orbit_wrong_ref},
there is essentially no spread in the determination of $J_{\phi}$. This is
expected since $J_{\phi}$ is independent of $z$ and is
thus unaffected by offsets in $z$, as discussed earlier. Indeed, the result we found in
Figure~\ref{fig:one_orbit_wrong_ref} for the thick-disk orbit holds for all
orbit types. This is also a demonstration of the robustness of the orbit
integration and action calculation methods we use.

The {\em upper right} panel of Figure~\ref{fig:many_orbit_wrong_ref} shows
that the fractional error in $J_z$ is more exaggerated for more planar
(disk-like) orbits. For the thin-disk orbit, a systematic offset of
$\sim15\,\pc$ in the $z$ coordinate already results in $25\%$ deviations in
the actions. The required offset for $25\%$ deviation is $\sim120\,\pc$ for
the thick-disk orbit. The halo orbit is relatively resistant to errors in the
midplane, with only $\sim15\%$ error in $J_z$ out to a midplane offset of
$500\,\pc$.

For the offset in the Solar radius ({\em bottom panels}), the error is largest
for $J_R$, with some deviations resulting in $J_{\phi}$ and relatively small
deviations in $J_z$. In the {\em bottom middle} and {\em bottom right} panels
all three lines nearly overlap.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{fig/schmactions_many_orbits.pdf}
\end{center}
\caption{We report one half the 95\uth{} minus 5\uth{} percentile of the error
in the measured action ($\Delta J_i$) from coordinate system errors for the
thin, thick, and halo orbits (Table~\ref{tab:orbits}). See discussion in the
text for the justification in using this to measure the magnitude of the
induced error. The {\em left}, {\em center}, and {\em right} panels show the
result for $J_R$, $J_{\phi}$, and $J_z$, respectively. The {\em upper} panels
consider an offset in $z$ and the {\em lower} panels consider an offset in $x$
(equivalently, an offset in the Solar radius). In some panels we also plot as
dashed lines the epicyclic prediction of the induced action error
(Equation~\eqref{eq:Ji_err_mosttime}). In the epicyclic approximation, a $z$
offset only induces an error in $J_z$ --- for all three orbits the epicyclic
approximation is a good description of the $J_z$ error. An $x$ offset induces
an error in $J_R$ and $J_{\phi}$. The error in $J_R$ is somewhat
well-described for the thin-disk and thick-disk orbits, and a poor description
for the halo orbit. For $J_{\phi}$, the epicyclic approximation is not a good
description for any orbit.}
\label{fig:many_orbit_wrong_ref}
\end{figure*}

@@

We compute the same midplane variation in Figure~\ref{fig:midplane_chervin},
but for the toy simulation of a Sagittarius encounter
\citep{2018MNRAS.481..286L}. Again we have subtracted a best fit curve of the
form $\text{A} \sin{\left(\phi + \text{B}\right)} + \text{C}$, with the values
for $\text{A}$ being $9.5$, $2.5$, $-212$, and $-394\,\pc$, for $\text{B}$
being $0.0013$, $0.0007$, $-0.63$, and $-1.00\,\text{rad}$ and $\text{C}$
being $2.6$, $-7.8$, $-65$, and $-53\,\pc$ chronologically for each of the four timesteps shown. In the {\em upper} two panels, we see that the midplane is
relatively flat in the inner galaxy, but additional encounters drive strong
midplane variation. In the {\em lower left} panel, we see a strong $m=2$ mode develop,
consistent with the $R=8\,\kpc$ panel of Figure~17 in
\citet{2018MNRAS.481..286L}.

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{fig/midplane_fit.pdf}
\end{center}
\caption{The local midplane determined at the fiducial Solar radius
($8.2\,\kpc$) for the three FIRE galaxies \mi{}, \mf{}, and \mm{} ({\em left},
{\em center}, and {\em right} panels) as a function of azimuthal angle. All calculations are done at
cosmological redshift $\z =0$. The local midplane is determined at a position
$\phi$ by taking the median height of all stars within $R=0.5\,\kpc$ and
$z=1\,\kpc$ (in cylindrical coordinates). The procedure is performed again
using the new height $10$ times to converge on the local midplane height. In
order to allow for the possibility that the fiducial galactocentric coordinate
system is incorrect, we subtract the best fit $A\sin{(\phi+B)}+C$ curve from
each panel --- this figure is reproduced with the original midplane
determination (i.e. before subtracting the best fit sine curve) in
Appendix~\ref{app:lsr}. We then bootstrap resample $1000$ times on all stars
within a $2\,\kpc$ height of the fiducial midplane to determine $1\,\sigma $
error bars, which we report as dashed lines.}
\label{fig:midplane}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=342.078286667pt]{fig/midplane_fit_chervinsim.pdf}
\end{center}
\caption{The local midplane determined at the fiducial Solar radius
($8.2\,\kpc$) for 4 different time steps from toy simulations
of a Sagittarius encounter with the Milky Way \citep{2018MNRAS.481..286L}. As
before, we have subtracted the best fit $A\sin{(\phi+B)}+C$ curve to account
for inaccuracies in the galactocentric coordinate system. Error bars are
calculated as in Figure~\ref{fig:midplane}. (\Gus{Check:}) The {\em upper}
panels show the midplane as a function of azimuth before the first encounter
near the Solar circle. As later encounters occur which bring the
Sagittarius-like object closer to the Solar circle, we can see that stronger
midplane variations result. In fact, one can see that an $m=2$ mode is excited
in the {\em lower left} panel --- this is consistent with Figure~17 of
\citet{2018MNRAS.481..286L}, which shows that an $m=2$ mode is excited (note
that $m=0$ and $m=1$ modes are stronger, but these are removed in our
sine-curve subtraction). The fact that the $t=6.9\,\Gyr$ panel looks
qualitatively similar to the panels from the FIRE simulations
(Figure~\ref{fig:midplane}) is evidence that the midplane variations are
produced in part by mergers.}
\label{fig:midplane_chervin}
\end{figure*}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{fig/orbits.pdf}
\end{center}
\caption{The three orbits presented in Table~\ref{tab:orbits} and considered
throughout the work. We plot the thin, thick, and halo orbits in the {\em
left}, {\em center}, and {\em right} columns, respectively. The {\em upper}
row shows a plot of $x$ vs. $y$ while the {\em lower} row shows $R$ vs. $z$.}
\label{fig:plot_orbits}
\end{figure*}

>>>>

We consider that the immortal observer makes a measurement every $\Myr$. At each measurement, we compute the actions as outlined above using the true potential but by specifiying the star's starting position using the systematically offset coordinate system.
Essentially we are shifting the
integrated orbit and then reintegrating at each point along the shifted orbit,
where the shifted orbit is the observed orbit. The erroneous computed actions
for each phase-space starting point are shown for the first $\Gyr$ of the
orbit in the upper panels
Figure~\ref{fig:one_orbit_wrong_ref}.\footnote{Occasionally the numerical
scheme fails and very large actions are reported by \texttt{gala}---we perform
a $5\sigma$ clip on each action to exclude such orbits, but this only excludes
a total of $5$ orbits out of the $1000$ considered for
Figure~\ref{fig:one_orbit_wrong_ref}. Some numerical artifacts remain, but the
vast majority of orbits are computed properly.}

@@

For a systematic offset in $z$ (upper panels), the
95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $2.2$ and
$6.2\,\actunit$ for $J_R$ and $J_z$, respectively. These are fractional errors
of $5.7\,\%$ and $86\,\%$, respectively. The error induced in $J_{\phi}$ is
negligible, as expected since $J_{\phi}$ only depends on the $x$- and
$y$-components of the position and velocity of the stars.\footnote{In practice,
however, $J_{\phi}$ is computed as part of the torus-fitting method.} It is worth
pointing out that a $100\,\pc$ error in an orbit with
$z_{\text{max}}=850\,\pc$ --- a $12\%$ error
--- induced an $86\%$ error in the computation of $J_z$.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{fig/schmactions_one_orbit.pdf}
\end{center}
\caption{The artificial phase-dependence in the observed actions induced by an
error in the Galactocentric coordinate system. We consider here the thick-disk
orbit, which has actions of $(J_R, J_{\phi}, J_z) = (37.9, -1520,
7.0)\,\actunit$ and $z_{\text{max}}=850\,\pc$ (see Table~\ref{tab:orbits}). We
integrate the orbit according to the procedure laid out in
Section~\ref{ssec:action_comp}, and which we plot in Appendix~\ref{app:orbits}. Then, we subtract $100\,\pc$ from the $z$
value (upper panels) or the $x$ value (lower panels) of each
position in the orbit, corresponding to an erroneous observer assuming a
midplane (upper) or solar radius (lower) that is off by
$100\,\pc$. We then allow an immortal observer to observe the orbit over
$1\,\Gyr$ and perform the same orbit integration procedure at each timestep,
and report the values of the actions. The computation of $J_{\phi}$ is
pristine to errors in $z$, with only numerical artifacts remaining. Only small
errors are induced in $J_R$, with the middle $90\%$ of values over the $\Gyr$
being within $\sim8\%$ of the true $J_R$. As expected, large errors are
induced in $J_z$ with a $100\,\pc$ offset in $z$, with the middle $90\%$ of
values being within $\sim43\%$ of the true $J_z$. The $x$ offset induces
uncertainties in $J_R$ and $J_{\phi}$ of $\sim21\%$ and $\sim3\%$,
respectively. A $\sim3\%$ error in $J_z$ is also induced.}
\label{fig:one_orbit_wrong_ref}
\end{figure*}

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{fig/schmactions_Jz_zerr_hist.pdf}
\end{center}
\caption{A histogram of the observed values in $J_z$ for the thick-disk orbit
assuming a $z$ offset of $100\,\pc$. One can see that if the observed $z$
values have a bias (from e.g. an incorrectly computed midplane), then the
induced error distribution in $J_z$ is decidedly non-Gaussian. Therefore, any
sort of error propagation must take this into account. Intuition for the shape
of each panel is given in the text. We also plot one half the 95\uth{}
percentile minus the 5\uth{} percentile of each distribution as a horizontal
arrow anchored on the true $J_z$ value. We call this $\Delta J_z$ and will use
it (along with the similarly defined $\Delta J_R$ and $\Delta J_{\phi}$) to
empirically describe the error distribution. We see that $\Delta J_z$ roughly
corresponds to the distance from the true $J_z$ value to one of the modes of
the error distribution.}
\label{fig:Jz_hist}
\end{figure}

In Figure~\ref{fig:Jz_hist} we plot a histogram of the observed values of
$J_z$ for the thick-disk orbit (top panel) and thin-disk orbit (bottom panel) assuming a $z$ offset of $100\,\pc$ (i.e. as in the upper
right panel of Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$
is plotted as a vertical dashed line. Here we see that the errors in $J_z$
induced by a systematic offset in $z$ are non-Gaussian and bimodal.
Furthermore, neither of the modes lie at zero. In other words, neither of the
modes of the distribution of observed $J_z$ values are centered on the true
$J_z$ (i.e. the $J_z$ computed with no offset in $z$). In the case of the
thin-disk orbit (bottom panel), we see that, in addition to the prior
complications, the error distribution is not even centered on null. This comes
about when the midplane error is $\gtrsim z_{\text{max}}$, where
$z_{\text{max}}$ is the maximum height of the orbit (equivalent to $A_z$ in
Section~\ref{ssec:epi_action}).

We now describe the intuition behind the shape of Figure~\ref{fig:Jz_hist}.
Consider first the thick-disk orbit (top panel), where the offset in $z$
is much less than $z_{\text{max}}$ of the orbit. The peaks in the distribution
correspond to the turning points of the orbit (or points of maximum vertical
amplitude), where $v_z \sim 0$ and where the star is on most of it orbit.
This is why the distribution is peaked on these values. For the thin-disk
orbit (bottom panel), the offset in $z$ is comparable to
$z_{\text{max}}$. Now, there will be some points in the orbit where $v_z = 0$
and $z=0$ (in the observed, erroneous coordinate system). At these points, the
computed $J_z$ will vanish. The asymmetry and systematic offset then comes
about because of the constraint that $J_z \geq 0$.\footnote{This is similar to
arguments in cosmology for why gravity produces non-Gaussianity in the density
field, since the density cannot become negative but it can grow arbitrarily
large.}

@@

The upper row of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread induced
in each action for an offset in the $z$-component (i.e. the Galactic
midplane). In the lower row we consider offsets in the $x$ component
(i.e. the Solar radius). The left, center, and right columns
show the fractional spread in the values computed for $J_z$, $J_{\phi}$, and
$J_R$, respectively. We compute the fractional spread $\Delta J_i/J_i$, where
$\Delta J_i$ is defined in the previous paragraph and $J_i$ is the action
value computed in the true coordinate system, over the course of the first
$\Gyr$.

In the upper middle panel of Figure~\ref{fig:many_orbit_wrong_ref},
there is essentially no spread in the determination of $J_{\phi}$. This is
expected since $J_{\phi}$ is independent of $z$ and is
thus unaffected by offsets in $z$, as discussed earlier. Indeed, the result we found in
Figure~\ref{fig:one_orbit_wrong_ref} for the thick-disk orbit holds for all
orbit types. This is also a demonstration of the robustness of the orbit
integration and action calculation methods we use.

The upper right panel of Figure~\ref{fig:many_orbit_wrong_ref} shows
that the fractional error in $J_z$ is more exaggerated for more planar
(disk-like) orbits. For the thin-disk orbit, a systematic offset of
$\sim15\,\pc$ in the $z$ coordinate already results in $25\%$ deviations in
the actions. The required offset for $25\%$ deviation is $\sim120\,\pc$ for
the thick-disk orbit. The halo orbit is relatively resistant to errors in the
midplane, with only $\sim15\%$ error in $J_z$ out to a midplane offset of
$500\,\pc$.

For the offset in the Solar radius (lower panels), the error is largest
for $J_R$, with some deviations resulting in $J_{\phi}$ and relatively small
deviations in $J_z$. In the lower middle and lower right panels
all three lines nearly overlap.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{fig/schmactions_many_orbits.pdf}
\end{center}
\caption{We report one half the 95\uth{} minus 5\uth{} percentile of the error
in the measured action ($\Delta J_i$) from coordinate system errors for the
thin, thick, and halo orbits (Table~\ref{tab:orbits}). See discussion in the
text for the justification in using this to measure the magnitude of the
induced error. The left, center, and right panels show the
result for $J_R$, $J_{\phi}$, and $J_z$, respectively. The upper panels
consider an offset in $z$ and the lower panels consider an offset in $x$
(equivalently, an offset in the Solar radius). In some panels we also plot as
dashed lines the epicyclic prediction of the induced action error
(Equation~\eqref{eq:Ji_err_mosttime}). In the epicyclic approximation, a $z$
offset only induces an error in $J_z$ --- for all three orbits the epicyclic
approximation is a good description of the $J_z$ error. An $x$ offset induces
an error in $J_R$ and $J_{\phi}$. The error in $J_R$ is somewhat
well-described for the thin-disk and thick-disk orbits, and a poor description
for the halo orbit. For $J_{\phi}$, the epicyclic approximation is not a good
description for any orbit.}
\label{fig:many_orbit_wrong_ref}
\end{figure*}

@@

We compute the same midplane variation in Figure~\ref{fig:midplane_chervin},
but for the toy simulation of a Sagittarius encounter
\citep{2018MNRAS.481..286L}. Again we have subtracted a best fit curve of the
form $\text{A} \sin{\left(\phi + \text{B}\right)} + \text{C}$, with the values
for $\text{A}$ being $9.5$, $2.5$, $-212$, and $-394\,\pc$, for $\text{B}$
being $0.0013$, $0.0007$, $-0.63$, and $-1.00\,\text{rad}$ and $\text{C}$
being $2.6$, $-7.8$, $-65$, and $-53\,\pc$ chronologically for each of the four timesteps shown. In the {\em upper} two panels, we see that the midplane is
relatively flat in the inner galaxy, but additional encounters drive strong
midplane variation. In the lower left panel, we see a strong $m=2$ mode develop,
consistent with the $R=8\,\kpc$ panel of Figure~17 in
\citet{2018MNRAS.481..286L}.

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{fig/midplane_fit.pdf}
\end{center}
\caption{The local midplane determined at the fiducial Solar radius
($8.2\,\kpc$) for the three FIRE galaxies \mi{}, \mf{}, and \mm{} (left,
center, and right panels) as a function of azimuthal angle. All calculations are done at
cosmological redshift $\z =0$. The local midplane is determined at a position
$\phi$ by taking the median height of all stars within $R=0.5\,\kpc$ and
$z=1\,\kpc$ (in cylindrical coordinates). The procedure is performed again
using the new height $10$ times to converge on the local midplane height. In
order to allow for the possibility that the fiducial galactocentric coordinate
system is incorrect, we subtract the best fit $A\sin{(\phi+B)}+C$ curve from
each panel --- this figure is reproduced with the original midplane
determination (i.e. before subtracting the best fit sine curve) in
Appendix~\ref{app:lsr}. We then bootstrap resample $1000$ times on all stars
within a $2\,\kpc$ height of the fiducial midplane to determine $1\,\sigma $
error bars, which we report as dashed lines.}
\label{fig:midplane}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=342.078286667pt]{fig/midplane_fit_chervinsim.pdf}
\end{center}
\caption{The local midplane determined at the fiducial Solar radius
($8.2\,\kpc$) for 4 different time steps from toy simulations
of a Sagittarius encounter with the Milky Way \citep{2018MNRAS.481..286L}. As
before, we have subtracted the best fit $A\sin{(\phi+B)}+C$ curve to account
for inaccuracies in the galactocentric coordinate system. Error bars are
calculated as in Figure~\ref{fig:midplane}. (\Gus{Check:}) The upper
panels show the midplane as a function of azimuth before the first encounter
near the Solar circle. As later encounters occur which bring the
Sagittarius-like object closer to the Solar circle, we can see that stronger
midplane variations result. In fact, one can see that an $m=2$ mode is excited
in the lower left panel at $t=6.0\,\Gyr$ --- this is consistent with Figure~17 of
\citet{2018MNRAS.481..286L}, which shows that an $m=2$ mode is excited (note
that $m=0$ and $m=1$ modes are stronger, but these are removed in our
sine-curve subtraction). The fact that the $t=6.9\,\Gyr$ panel looks
qualitatively similar to the panels from the FIRE simulations
(Figure~\ref{fig:midplane}) is evidence that the midplane variations are
produced in part by mergers.}
\label{fig:midplane_chervin}
\end{figure*}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{fig/orbits.pdf}
\end{center}
\caption{The three orbits presented in Table~\ref{tab:orbits} and considered
throughout the work. We plot the thin-disk, thick-disk, and halo orbits in the left, center, and right columns, respectively. The upper
row shows a plot of $x$ vs. $y$ while the lower row shows $R$ vs. $z$.}
\label{fig:plot_orbits}
\end{figure*}
